Simulating High-Dimensional Multivariate Data

### 2021-02-23 010-background.Rmd
Polishing for CSDA submission, addressing review criteria

```{r ch010-realDataTab, echo = FALSE, eval=TRUE, message=FALSE, warning=FALSE}
## Describe and plot real data
if ( !file.exists( './data/exampleGenes.rds' ) )  {
    set.seed(01122021)
    exampleGenes <- colnames(brca)[sort(sample(ncol(brca), 3))]
    saveRDS(exampleGenes, file = './data/exampleGenes.rds')
}
exampleGenes <- readRDS('./data/exampleGenes.rds')
## Add row names per WWP comment
small <-  brca %>% select( exampleGenes ) %>% as.data.frame
numPatients <- 5
smallForTable <- head( small, n = numPatients )
smallForTable$PatientID <- head( readLines('./data/examplePatients.txt'), n = numPatients )
smallForTable <- smallForTable[ , c(4, 1:3 ) ]
smallForTable[ , 2:4 ] <- round( smallForTable[ , 2:4 ] )
knitr::kable(smallForTable ,
      booktabs = TRUE,
      caption = "mRNA expression for three selected high-expressing genes (*RPL5*, *TKNIP*, and *VIM*) for the first five patients in the TCGA BRCA data set."
)
```

>>>>

```{r ch010-realDataTab, echo = FALSE, eval=TRUE, message=FALSE, warning=FALSE}
## Describe and plot real data
if ( !file.exists( './data/exampleGenes.rds' ) )  {
    set.seed(01122021)
    exampleGenes <- colnames(brca)[sort(sample(ncol(brca), 3))]
    saveRDS(exampleGenes, file = './data/exampleGenes.rds')
}
exampleGenes <- readRDS('./data/exampleGenes.rds')
## Add row names per WWP comment
small <-  brca %>% select( exampleGenes ) %>% as.data.frame
numPatients <- 5
smallForTable <- head( small, n = numPatients )
smallForTable$PatientID <- head( readLines('./data/examplePatients.txt'), n = numPatients )
smallForTable <- smallForTable[ , c(4, 1:3 ) ]
smallForTable[ , 2:4 ] <- round( smallForTable[ , 2:4 ] )
knitr::kable(smallForTable ,
             booktabs = TRUE,
             row.names = F,
      caption = paste0( "mRNA expression for three selected high-expressing genes, ", paste( names(smallForTable)[-1], collapse = ", " ), ", for the first five patients in the TCGA BRCA data set.")
)
```

### 2021-02-23 030-bigsimr-package.Rmd
Polishing for CSDA submission, addressing review criteria

The `bigsimr` package is a high-performance implementation of our proposed random vector generation algorithm and associated functions (see Section \@ref(algorithms)).
When designing `bigsimr`, we aimed to conveniently provide parallelized computation, through multi-core and GPU acceleration, while allowing advanced users to customize and automate workflows.

@@

```{r ch030-basic-example-setup, echo=TRUE, message=FALSE, cache=F}
library(bigsimr)
library(tidyverse)
library(patchwork)
```

@@

```{r ch030-aq-cor, echo=TRUE}
type <- 'spearman'
(R_S <- cor_fast(df, method = "spearman"))
```

@@

```{r ch030-cor-bounds, echo=TRUE}
cor_bounds(margins = margins, method = type)
```

>>>>

The following descrption is directed towards towards `R` users via `bigsimr` R package.
This package provides an interface to the native code written in Julia, registered as the `Bigsimr`Julia package PROVIDE LINK.
In addition to the native Julia `Bigsimr` package and `R` interface `bigsimr`, we also provide a python interface `PyBigsimr` at URL that interfaces with the Julia `Bigsimr` package.
The Julia implementation provides a high-performance implementation of our proposed random vector generation algorithm and associated functions (see Section \@ref(algorithms)).
When designing `bigsimr`, we aimed to conveniently provide parallelized computation, through multi-core and GPU acceleration, while allowing advanced users to customize and automate workflows.

@@

```{r ch030-basic-example-setup, echo=TRUE, message=FALSE, cache=F}
## devtools::install_github("SchisslerGroup/bigsimr", ref="dev-julia")
library(bigsimr)
Sys.setenv(JULIA_NUM_THREADS=parallel::detectCores())
## Either let JuliaCall find your installation of Julia or set manually
bs <- bigsimr_setup(JULIA_HOME = "/usr/local/bin/")
dist <- distributions_setup() # Julia distributions
library(tidyverse)
library(patchwork)
```

@@

```{r ch030-aq-cor, echo=TRUE}
type <- 'spearman'
(R_S <- bs$cor(as.matrix(df), bs$Spearman))
```

@@

```{r ch030-cor-bounds, echo=TRUE}
bounds <- bs$cor_bounds(D)
bounds$lower
```

### 2021-02-23 040-monte-carlo-evaluation.Rmd
Polishing for CSDA submission, addressing review criteria

```{r ch040-LoadLib040, echo = FALSE, eval = TRUE, message=FALSE, warning=FALSE, results = "hide"}
## devtools::install_github("SchisslerGroup/bigsimr", ref = 'develop')
## knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
# reticulate::use_condaenv("bigsimr")
library(bigsimr)
library(tidyverse)
cores <- as.integer( parallel::detectCores() - 1 )
## cores <- as.integer( 1 )
set.seed(10162020)
```

@@

>>>>

```{r ch040-LoadLib040, echo = FALSE, eval = FALSE, message=FALSE, warning=FALSE, results = "hide"}
## devtools::install_github("SchisslerGroup/bigsimr", ref="dev-julia")
library(bigsimr)
Sys.setenv(JULIA_NUM_THREADS=parallel::detectCores())
## Either let JuliaCall find your installation of Julia or set manually
bs <- bigsimr_setup(JULIA_HOME = "/usr/local/bin/")
dist <- distributions_setup() # Julia distributions
library(tidyverse)
cores <- as.integer( parallel::detectCores() - 1 )
## cores <- as.integer( 1 )
set.seed(10162020)
```

@@

*Limitations, conclusions, and recommendations.*
In the bivariate studies, we chose arbitrary simulation parameters for three distributions, moving from the Gaussian to the discrete and non-normal, multivariate negative binomial.
Under these conditions, the simulated random vectors sample the desired bivariate distribution across the entire range of pairwise correlations for the three dependency measures.
The simulation results could differ for other choices of simulation settings.
Specifying extreme correlations near the boundary or Frechet bounds could result in poor simulation performance.
Fortunately, it is straightforward to evaluate simulation performance by using strategies similar to those completed above.
We expect our random vector generation to perform well for the vast majority of NORTA-feasible correlation matrices, but advise to check the performance before making inferences/further analyses.

### 2021-02-23 050-examples.Rmd
2nd draft. Insert Tom / Ania bivariate example
Polishing for CSDA submission, addressing review criteria

```{r ch050-cor-check, echo=TRUE, eval=TRUE, cache = FALSE}
## 1. Mapping step first
R_X <- cor_convert(rho = R_S, from = type, to = 'pearson')
## 2a. Check admissibility
is_valid_correlation <- function(rho) {
  all(diag(rho) == 1) && matrixcalc::is.positive.semi.definite(rho)
}
is_valid_correlation( R_X )

@@

>>>>

```{r ch050-cor-check, echo=TRUE, eval=TRUE, cache = TRUE}
## 1. Mapping step first
R_X <- cor_convert(rho = R_S, from = type, to = 'pearson')
## 2a. Check admissibility
is_valid_correlation <- function(rho) {
  all(diag(rho) == 1) && matrixcalc::is.positive.semi.definite(rho)
}
is_valid_correlation( R_X )

@@

*Limitations, conclusions, and recommendations.*
The results show good simulation performance for our choice of parameters settings.
Our settings were motivated by modeling high-expressing genes from TCGA BRCA data set.
In general, the ability to match marginal and dependence parameters depends on the particular joint probability model.
We recommend to evaluate and tune your simulation until you can be assured of the accuracy.

### 2021-02-23 index.Rmd
Polishing for CSDA submission, addressing review criteria

```{r indexSetup, echo = FALSE, eval = TRUE, message=FALSE, warning=FALSE, results = "hide", cache=F}
library(bigsimr)
library(tidyverse)
library(knitr)
knitr::opts_chunk$set(eval = TRUE, echo = FALSE, cache = TRUE)
library(rmarkdown)
library(bookdown)
## setwd('~/article_bigsimr')
set.seed(09212020)
```

>>>>

```{r indexSetup, echo = FALSE, eval = TRUE, message=FALSE, warning=FALSE, results = "hide", cache=F}
## devtools::install_github("SchisslerGroup/bigsimr", ref="dev-julia")
## library(bigsimr)
## Sys.setenv(JULIA_NUM_THREADS=parallel::detectCores())
## Either let JuliaCall find your installation of Julia or set manually
## bs <- bigsimr_setup(JULIA_HOME = "/usr/local/bin/")
## dist <- distributions_setup() # Julia distributions
library(tidyverse)
library(knitr)
library(printr)
knitr::opts_chunk$set(eval = TRUE, echo = FALSE, cache = TRUE)
library(rmarkdown)
library(bookdown)
## setwd('~/article_bigsimr')
set.seed(09212020)
```
